---
title: "covid_statespend_covariates_v2"
author: "Denise Upton"
date: "2024-02-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# set up
library(readr)
library(readxl)
library(lubridate)
library(tidyverse)
library(janitor)
library(DT)
```

## State Spending

* Housing, Health, & Public Welfare spending as a fraction of total spending
* Geolevel: state
* Frequency: yearly
* Source: [Urban Institute State Finance Data](https://state-local-finance-data.taxpolicycenter.org/pages.cfm)

```{r}
state_spending <- read_excel("~/Documents/Coding Projects/quant_capstone/data_raw/state_spending_urban.xlsx",
                           range = "A3:E207") %>% 
  clean_names("snake") %>%
  filter(state != "United States") %>% 
  left_join(read_excel("~/Documents/Coding Projects/quant_capstone/data_raw/state_abbrev_crosswalk.xlsx"),
            by = "state") %>% 
  rename(state_full = state,
         state = postal) %>% 
  mutate(state_fips = str_pad(as.character(state_fips), width = 2, side = "left", pad = "0"))

state_spending2 <-
  state_spending %>%
  rename(health_spending=e052_health_hosp_dir_exp, housing_spending=e074_hous_com_direct_exp, welfare_spending=e090_public_welf_direct_exp) %>%
  group_by(state, state_fips) %>%
  summarize(health_spending = mean(health_spending), housing_spending = mean(housing_spending), welfare_spending = mean(welfare_spending))
```

## COVID-19 Incidence Rate

* Count of COVID-19 infections per 100,000
* Geolevel: county
* Frequency: monthly
* Source: [CDC U.S. State and Territorial Public Mask Mandates From April 10, 2020 through July 20, 2021 by County by Day](https://data.cdc.gov/Public-Health-Surveillance/Weekly-COVID-19-County-Level-of-Community-Transmis/jgk8-6dpn/about_data)

```{r}

#read county-level data
county_covid_rates <- read_csv("~/Google Drive/My Drive/_Wagner/2024-01/Capstone/Data/Weekly_COVID-19_County_Level_of_Community_Transmission_Historical_Changes_-_ARCHIVED_20240213.csv") %>%
  clean_names("snake") %>%
  rename("covid_case_rate" = "cases_per_100k_7_day_count_change") %>% 
  mutate(date = as.Date(date, format = '%m/%d/%Y'))


#preview data
county_covid_rates %>%
  group_by(date) %>% 
  tally() %>% 
  arrange(date) %>% 
  datatable()

#aggregate to month-year average
county_covid_rates2 <-
  county_covid_rates %>%
  arrange(date) %>% 
  filter(date >= "2020-07-01") %>%
  mutate(ym = format(date, '%Y-%m')) %>% 
  group_by(ym, fips_code) %>% 
  summarize(monthly_covid_rate = mean(covid_case_rate)) %>% 
  separate(fips_code, into = c("state_fips", "county_fips"), sep = +2)

head(county_covid_rates2) %>% datatable()

#average county rate across study period
county_covid_rates3 <-
  county_covid_rates2 %>%
  group_by(state_fips, county_fips) %>%
  summarize(study_covid_rate = mean(monthly_covid_rate))

head(county_covid_rates3) %>% datatable()

#average county rate across pre-treatment period
county_covid_rates4 <-
  county_covid_rates2 %>%
  arrange(ym) %>% 
  filter(ym <= "2021-09") %>%
  group_by(state_fips, county_fips) %>%
  summarize(pretreat_covid_rate = mean(monthly_covid_rate))

head(county_covid_rates4) %>% datatable()

```

* convert  

## Mask Mandates

* Percent of study period mask mandate is in place
* Geolevel: county
* Frequency: 3xmonthly
* Source: [CDC U.S. State and Territorial Public Mask Mandates From April 10, 2020 through July 20, 2021 by County by Day](https://data.cdc.gov/Policy-Surveillance/U-S-State-and-Territorial-Public-Mask-Mandates-Fro/62d6-pm5i/about_data)
* Notes: masks_order_code 1=yes; 2=no

make no=0
sum yes/# dates over study period (07/01/2020+)

```{r}

#read county-level data
county_mask_mandates <- read_csv("~/Google Drive/My Drive/_Wagner/2024-01/Capstone/Data/U.S._State_and_Territorial_Public_Mask_Mandates_From_April_10__2020_through_July_20__2021_by_County_by_Day_20240217.csv") %>%
  clean_names("snake")

#preview data
county_mask_mandates %>%
  group_by(date) %>% 
  tally() %>% 
  arrange(date) %>% 
  datatable()

#aggregate data to month-year
county_mask_mandates2 <-
  county_mask_mandates %>%
  mutate(masks_order_code = replace(masks_order_code, masks_order_code == 2, "0")) %>%
  filter(date >= "2020-07-01") %>%
  mutate(ym = format(date, '%Y-%m')) %>% 
  group_by(ym, fips_code) %>% 
  summarize(masks_order_code = max(masks_order_code)) %>%
  separate(fips_code, into = c("state_fips", "county_fips"), sep = +2)


head(county_mask_mandates2) %>% datatable()

#percent of study period with mask mandate
county_mask_mandates3 <-
  county_mask_mandates2 %>%
  mutate(masks_order_code = as.numeric(masks_order_code)) %>%
  group_by(state_fips, county_fips) %>%
  summarize(study_maskmandate_pct = mean(masks_order_code))
  
head(county_mask_mandates3) %>% datatable()

#percent of pre-treatment period with mask mandate
county_mask_mandates4 <-
  county_mask_mandates2 %>%
  mutate(masks_order_code = as.numeric(masks_order_code)) %>%
  filter(ym <= "2021-09") %>%
  group_by(state_fips, county_fips) %>%
  summarize(pretreat_maskmandate_pct = mean(masks_order_code))

```


## Join variables and export

```{r}

covariates1 <- full_join(state_spending2, county_covid_rates2, by = c("state_fips"))
covariates2 <- full_join(covariates1, county_covid_rates3, by = c("county_fips", "state_fips"))
covariates3 <- full_join(covariates2, county_covid_rates4, by = c("county_fips", "state_fips"))
covariates4 <- full_join(covariates3, county_mask_mandates2, by = c("ym", "county_fips", "state_fips"))
covariates5 <- full_join(covariates4, county_mask_mandates3, by = c("county_fips", "state_fips"))
covariates6 <- full_join(covariates5, county_mask_mandates4, by = c("county_fips", "state_fips"))



```