---
title: "COVID and State Gov Covariates"
author: "Camille Preel-Dumas"
date: 'Updated: `r format(Sys.time(), ''%a, %b %d, %Y at %I:%M %p'')`'
output:
  pdf_document:
    toc: true
    toc_depth: '3'
  html_document:
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: false
    df_print: kable
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# set up
library(readr)
library(readxl)
library(lubridate)
library(tidyverse)
library(janitor)
library(DT)
```

Note: change files paths to your own to read data

<br>

# State Spending 

I downloaded Housing, Health, & Public Welfare spending as a fraction of total spending, but other categories are available. 

* Geolevel: state
* Frequency: yearly
* Source: [Urban Institute State Finance Data](https://state-local-finance-data.taxpolicycenter.org/pages.cfm)

This dataset is ready to merge to our main BLS data.
```{r}
state_spending <- read_excel("/Users/cpreeldumas/Documents/GitHub/quant_capstone/data_raw/state_spending_urban.xlsx",
                           range = "A3:E207") %>% 
  clean_names("snake") %>%
  filter(state != "United States") %>% 
  left_join(read_excel("/Users/cpreeldumas/Documents/GitHub/quant_capstone/data_raw/state_abbrev_crosswalk.xlsx"),
            by = "state") %>% 
  rename(state_full = state,
         state = postal) %>% 
  mutate(state_fips = str_pad(as.character(state_fips), width = 2, side = "left", pad = "0"))

head(state_spending) %>% datatable()
```

<br> 

# Covid Cases

Issues:
1. [County-level case data](https://data.cdc.gov/Public-Health-Surveillance/Weekly-COVID-19-County-Level-of-Community-Transmis/dt66-w6m6/about_data) start in 2022 even though the dataset page says the timeline is 2020-01-22/2023-05-10 ?
2. I did find [state-level cases](https://data.cdc.gov/Case-Surveillance/Weekly-United-States-COVID-19-Cases-and-Deaths-by-/pwn4-m3yp/about_data) going back to 2020, if we're fine to proceed with state-level only

https://data.cdc.gov/Public-Health-Surveillance/Weekly-COVID-19-County-Level-of-Community-Transmis/jgk8-6dpn/about_data

#### County-level Cases #1:
```{r}

#read county-level
county_covid_cases <- read_csv("/Users/cpreeldumas/Documents/GitHub/quant_capstone/data_raw/Weekly_COVID-19_County_Level_of_Community_Transmission_as_Originally_Posted_-_ARCHIVED_20240130.csv") %>% 
  clean_names("snake")

#sort by report date to see earliest
county_covid_cases %>% 
  mutate(report_date = mdy(report_date)) %>% 
  group_by(report_date) %>% 
  tally() %>% 
  arrange(report_date) %>% 
  datatable()
```

#### County-level Cases #2:
```{r}

#read county-level
county_covid_cases2 <- read_csv("~/Google Drive/My Drive/_Wagner/2024-01/Capstone/Data/Weekly_COVID-19_County_Level_of_Community_Transmission_Historical_Changes_-_ARCHIVED_20240213.csv") %>% 
  clean_names("snake")

#sort by report date to see earliest
county_covid_cases2 %>% 
  mutate(date = mdy(date)) %>% 
  group_by(date) %>% 
  tally() %>% 
  arrange(date) %>% 
  datatable()
```
<br> 

#### State-level Cases:
```{r}
#state-level
state_covid_cases <- read_csv("/Users/cpreeldumas/Documents/GitHub/quant_capstone/data_raw/Weekly_United_States_COVID-19_Cases_and_Deaths_by_State_-_ARCHIVED_20240130.csv")

#sort by report date to see earliest
state_covid_cases %>% 
  mutate(start_date = mdy(start_date)) %>% 
  group_by(start_date) %>% 
  tally() %>% 
  arrange(start_date) %>% 
  datatable() 
```

<br> 

# Covid Deaths

This is total [state level deaths](https://data.cdc.gov/NCHS/Provisional-COVID-19-Death-Counts-in-the-United-St/kn79-hsxy/about_data) from 2020-2023 (not monthly or yearly). We might want to check if there is more granular data (county and/or monthly).

We also should decide whether deaths or cases make more sense here. Deaths might be more reliably reported and less subject to racial inequalities in case testing/reporting, but they're also a more rare/extreme outcome.

```{r}
deaths <- read_csv("/Users/cpreeldumas/Documents/GitHub/quant_capstone/data_raw/Provisional_COVID-19_Death_Counts_in_the_United_States_by_County_20240130.csv") %>% 
  clean_names("snake")
```

** Use total deaths (not just COVID)? **

[Weekly Provisional Counts of Deaths by State and Select Causes, 2020-2023](https://data.cdc.gov/NCHS/Weekly-Provisional-Counts-of-Deaths-by-State-and-S/muzy-jte6/about_data)



<br> 

# Mask Mandates

This is [daily mask mandate data](https://data.cdc.gov/Policy-Surveillance/U-S-State-and-Territorial-Public-Mask-Mandates-Fro/62d6-pm5i/about_data) So we need to summarize to monthly. 

Note that it is too large to push to github, so it must be read from a local path.

* columns with date, state, binary for whether order in place?, start date, end date

```{r}

```

<br> 

# State party affiliation

I haven't found a ready-made, national dataset for this yet. The closest I've found are yearly PDFs from the [National Conference of State Legislatures.](https://www.ncsl.org/about-state-legislatures/state-partisan-composition), which we would have to scrape. I think this covariate is less of a priority however.
```{r}

```

<br> 

# Stay at home orders

There is a [ready-made dataset](https://www.kff.org/policy-watch/stay-at-home-orders-to-fight-covid19/) with the start date of these orders by state, but it does not have the lift date. There are a few news sources with lift dates, but it seems like we would have to compile the dataset ourselves. This seems low priority though since these orders would only last a couple months into 2020.

News sources on lifts:

* https://www.usatoday.com/storytelling/coronavirus-reopening-america-map/
* https://abcnews.go.com/US/list-states-stay-home-order-lifts/story?id=70317035
* https://www.nytimes.com/interactive/2020/us/coronavirus-stay-at-home-order.html

[Daily data set with orders by county](https://data.cdc.gov/Policy-Surveillance/U-S-State-and-Territorial-Stay-At-Home-Orders-Marc/y2iy-8irm/about_data)

```{r}

```

<br> 

# School closures

Haven't investigated this one much, I've just come across [this news source](https://www.edweek.org/leadership/map-coronavirus-and-school-closures-in-2019-2020/2020/03) with state level closure date, and maybe [this page](https://www.edweek.org/leadership/map-where-are-schools-closed/2020/07) has lift dates.
```{r}

```

<br> 

