---
title: "DiD Initial Exploration"
author: "Camille Preel-Dumas"
date: "Updated: `r format(Sys.time(), '%a, %b %d, %Y at %I:%M %p')`"
output: 
  html_document:
    highlight: tango
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: false
    df_print: kable
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# set up
library(readr)
library(lubridate)
library(tidyverse)
library(janitor)
library(did)
library(naniar)

#import data
bls_data <- read_csv("/Users/cpreeldumas/Documents/GitHub/quant_capstone/data_clean/BLS_county_employment_clean.csv") %>% 
  clean_names("snake") #convert column names to snake case
```

# Initial Exploration Phase

This is a first pass at using the CS-DID package. Next steps include:

1. Adding covariates
2. Adding anticipatory effects (if any) - might need to test the other deadlines that got pushed (to test how solid the anticipation is). Fir ex: moratorium is initially set to expire in March, then it's pushed in February a few months, we should already be seeing effect in Jan and Feb.
3. Testing whether seasonality poses a threat
4. Confirming the parallel trends assumption

# Methodology decisions

1. Study period: July 2020 - April 2023 (expiration of MA moratorium)

# Note on Covariates

* The did package requires that covariates be time-invariant. For time varying covariates, the did package sets the value of the covariate to be equal to the value of the covariate in the “base period” where, 
  * in post-treatment periods the base period is the period immediately before observations in a particular group become treated (when there is anticipation, it is before anticipation effects start too), 
  * and in pre-treatment periods the base period is the period right before the current period. 

# Prep data for CS-DID
```{r, include = FALSE}
#missingness
miss_var_summary(bls_data)

#why are these state missing state_start? 
bls_data %>% 
  filter(is.na(state_start)) %>% 
  group_by(state) %>% 
  tally()

#prep dataset for CS-DID package
bls_data_did <- bls_data %>% 
  mutate(idname = as.numeric(fips)) %>% 
  # limiting dataset to July 2020 - April 2023 (expiration of MA moratorium)
  filter(date >= "2020-07-01" & date <= "2023-04-01") %>% 
  # making numeric version of data for DID package
  mutate(date_num = paste0(year, str_pad(month,
                                         width = 2,
                                         side = "left",
                                         pad = "0"))) %>% 
  mutate(across(c(state_start, state_end, federal_start, federal_end), mdy)) %>% 
  #making a numeric version of dates for first treat column
  mutate(across(
    .cols = c(state_start, state_end, federal_start, federal_end), 
    .fns = ~ paste0(year(.x), str_pad(month(.x),
                                      width = 2,
                                      side = "left",
                                      pad = "0")),
    .names = "{.col}_num")) %>% 
  mutate(across(ends_with("_num"), as.numeric)) %>% 
  mutate(first_treat = case_when(
    state == "MA" ~ 0, #never-treated group
    treatment == 1 ~ federal_end_num,
    treatment == 0 ~ state_end_num)) 

bls_data_did %>% 
  group_by(state, treatment, federal_end, state_end, first_treat) %>% 
  tally()

bls_data_did %>% 
  group_by(date, date_num) %>% 
  tally() %>% 
  arrange(date_num)

# checking the coding of treatment v. control
bls_data_did %>% 
  group_by(treatment, federal_end, state_end, first_treat) %>% 
  tally()

# DC is a much smaller group, saving a version without
bls_data_did %>% 
  group_by(first_treat) %>% 
  tally()

bls_data_did_nodc <- bls_data_did %>% 
  filter(state != "DC")
```

# CS-DID

idname = fips
tname = date
yname = unemployment_rate
gname = first_treat
xformla = ~1 

### No covariates

#### Group-time average treatment effects
```{r}
attgt_nocovar <- att_gt(yname = "unemployment_rate",
                   gname = "first_treat",
                   idname = "idname",
                   tname = "date_num",
                   xformla = ~1,
                   data = bls_data_did_nodc)

summary(attgt_nocovar)
ggdid(attgt_nocovar, ylim = c(-.3, .3), xgap = 12)
```

#### Aggregate 
```{r}
# aggregate the group-time average treatment effects
attgt_nocovar_dyn <- aggte(attgt_nocovar, type = "dynamic")
summary(attgt_nocovar_dyn)

ggdid(attgt_nocovar_dyn, ylim = c(-.3, .3), xgap = 10)
```
# Issues so far

1. Large standard errors
